---
title: "Data Analysis Assignment 5"
author: "Himangshu Raj Bhantana"
date: "11/11/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Part 1. Asthma Patients in California

**Are the covariates in this data balanced between the two groups? If no, which covariates are not? How did you assess balance?**

Covariates isn't balanced between two groups, there were 8 unbalanced covariate which can been seen below. We use the ASD (absolute standardized difference) to assess the balance, where the absolute standardized difference value is greater than 0.1.

```{r, results= "asis", echo=FALSE, message=FALSE, warning=FALSE}
library(xtable)
library(MatchIt) #for propensity score matching
library(cobalt)
library(ggplot2)
library(randomForest)
library(dplyr)
Asthma <- read.csv("~/Downloads/Asthma.txt", sep="")
#str(Asthma)
#head(Asthma)
#dim(Asthma)
#summary(Asthma)


Asthma$i_race<- factor(Asthma$i_race)
Asthma$i_educ<- factor(Asthma$i_educ)
Asthma$i_insu<- factor(Asthma$i_insu)
Asthma$i_drug<- factor(Asthma$i_drug)
Asthma$i_seve<- factor(Asthma$i_seve)

Asthma$i_sex <- relevel(factor(Asthma$i_sex), ref = 1)
Asthma$i_educ <- relevel(factor(Asthma$i_educ), ref = 5)
Asthma$i_seve <- relevel(factor(Asthma$i_seve), ref = 3)

Asthma$pgy<- 1
Asthma$pgy[Asthma$pg == 2] <- 0
asthma$pgy <- as.factor(asthma$pg)

Asthma$com_t <- Asthma$com_t - mean(Asthma$com_t)
Asthma$pcs_sd <- Asthma$pcs_sd - mean(Asthma$pcs_sd)
Asthma$mcs_sd <- Asthma$mcs_sd - mean(Asthma$mcs_sd)

bt=bal.tab(list(treat=Asthma$pgy,covs=Asthma,estimand="ATE"))
ddg<-bt$Balance %>% filter(abs(Diff.Un) > 0.1)
love.plot(list(treat=Asthma$pgy,covs=Asthma,estimand="ATE"),stars = "std")
xtable(ddg, type='latex', title='Baseline Model Summary', header= FALSE, digits=2, no.space = TRUE)
```
## Part2 

## Question 1
**Are there any observations with an estimated propensity score e that is out of the range of e in the other group? If there are only a few such outliers (less than 5), keep them; If many, discard them and report the number of the discarded observations.**

There are observations with an estimated propensity score e that is outside the range of e in the other group; there were more than 5, therefore we chose to discard them; the number of observations discarded was 48, with one group having a number of 40 and another having a number of 8.

Using one-to-one, nearest neighbor matching on the estimated propensity scores, check balance again. Are the covariates balanced now? If no, which ones are not?

Using one-to-one matching and balancing still gave us the covariates unbalanced 

